---
title: "Medflex"
author: "Sol Libesman"
output: 
  github_document:
---


```{r loading packages,include=FALSE}
library(tidyverse)
library(lubridate)
library(broom)
library(janitor)
library(flextable)
library(readxl)
#for negative binomial models it affects a few functions so using MASS:: instead
#library(MASS)
#for mediation analysis
library(mediation)
library(medflex)
# for viewing models, one i think was for astrix based on pvlaue
library(pacman)
library(gtools)
library(cowplot)
library(flextable)
library(gtsummary)
library(knitr)
library(kableExtra)

```


```{r chunk_options, include=FALSE}

opts_chunk$set(echo = TRUE,
               comment = "#>",
               out.width = "100%")
```





```{r, echo=F}
load("Y:\\Systematic Reviews\\TITANS/Data_checking_TITANS/TITANS_final_DF.rda")
```


```{r, echo=F, warning =F}
#knit(input="Analysis_medflex_for_Git2.md", output = "TITANS_medlfex.md") 

final_df <- titans_final_df


final_df <- final_df %>% filter(!is.na(treatment_cat))

#formating
final_df <- final_df %>% mutate(across(c(treatment_cat1,multiple,rc_transfusion_titans, mech_vent_combined), as.factor),
                                 across(c(transfusions_n_combined, cumalitive_blood_vol_sampled, arterial_lines, calculated_total_arterial_line_time_hr, blood_pressure_low_mean), function(x) as.numeric(as.character(x)) )
)



# making infants that were no to 'any transfusion', and NA for 'trans count' now 0
final_df$transfusions_n_combined <- ifelse((is.na(final_df$transfusions_n_combined)|final_df$transfusions_n_combined=="0") & final_df$rc_transfusion_titans=="0", 0,final_df$transfusions_n_combined)



final_df$transfusions_n_combined <- ifelse(final_df$transfusions_n_combined==0 & final_df$rc_transfusion_titans=="1", NA,final_df$transfusions_n_combined)


#logic check of coding for transfusion vol
final_df$calculated_total_rc_transfusions_vol1 <- ifelse(is.na(final_df$calculated_total_rc_transfusions_vol1) & final_df$rc_transfusion_titans==0, 0,final_df$calculated_total_rc_transfusions_vol1)


final_df$calculated_total_rc_transfusions_vol1 <- ifelse(final_df$calculated_total_rc_transfusions_vol1==0 & final_df$rc_transfusion_titans==1, NA,final_df$calculated_total_rc_transfusions_vol1)


# making infants that were yes to 'any transfusion' and 'trans vol' = 0 into NA
final_df$calculated_total_rc_transfusions_vol <- ifelse(final_df$rc_transfusion_titans=="1"&final_df$calculated_total_rc_transfusions_vol==0, NA,final_df$calculated_total_rc_transfusions_vol)



final_df$calculated_total_rc_transfusions_vol <- coalesce(final_df$calculated_total_rc_transfusions_vol.x, final_df$calculated_total_rc_transfusions_vol.y)

# making infants that were no to receive 'any transfusion', and NA for 'trans vol' now 0
final_df$calculated_total_rc_transfusions_vol <- ifelse(is.na(final_df$calculated_total_rc_transfusions_vol) & final_df$rc_transfusion_titans==0, 0,final_df$calculated_total_rc_transfusions_vol)

final_df$calculated_total_rc_transfusions_vol_scaled <- final_df$calculated_total_rc_transfusions_vol/100 



#filter out hypoplastic anaemia and aplastic anaemia
final_df <- final_df %>% filter((hypoplastic_anaemia!="1"|is.na(hypoplastic_anaemia)) &(aplastic_anaemia!="1"|is.na(aplastic_anaemia)) )


#scalling because numbers are so large
final_df$cumalitive_blood_vol_sampled_scaled <- final_df$cumalitive_blood_vol_sampled/1000

#removing outliers
 final_df$calculated_total_arterial_line_time_hr <- ifelse(final_df$calculated_total_arterial_line_time_hr<0|final_df$calculated_total_arterial_line_time_hr>2000, NA, final_df$calculated_total_arterial_line_time_hr)
  

```

# Checking associations

### Between treatment and mediators


```{r, warning =F}
treatmentmodel <- glm(treatment_cat1  ~ hct_week1_peak+cumalitive_blood_vol_sampled_scaled +arterial_lines+mech_vent_combined,  family =  binomial("logit"), data = final_df)

 glm(treatment_cat1  ~ hct_week1_peak+cumalitive_blood_vol_sampled_scaled +arterial_lines+mech_vent_combined,  family =  binomial("logit"), data = final_df) %>% tbl_regression(exponentiate = TRUE)


```


### global model

examining the relationship between treatment and transfusions (any) adjusting for all covariates


```{r, warning =F}



adjustedfullmodel <- glm(rc_transfusion_titans ~ treatment_cat1+ hct_week1_peak+cumalitive_blood_vol_sampled_scaled +arterial_lines+mech_vent_combined +GA_weeks_and_days_integer + multiple,  family =  binomial("logit"), data = final_df)


glm(rc_transfusion_titans ~ treatment_cat1+ hct_week1_peak+cumalitive_blood_vol_sampled_scaled +arterial_lines+mech_vent_combined +GA_weeks_and_days_integer + multiple,  family =  binomial("logit"), data = final_df) %>% tbl_regression(exponentiate = TRUE)


```




```{r, warning =F}
unadjustedmodel <- glm(rc_transfusion_titans ~ treatment_cat1,  family =  binomial("logit"), data = final_df) %>%
  tbl_regression(exponentiate = TRUE,
                 show_single_row="treatment_cat1",
                 include = c("treatment_cat1"))
```


```{r, warning =F}
adjusted_basedlinedmodel <- glm(rc_transfusion_titans ~ treatment_cat1+GA_weeks_and_days_integer + multiple,  family =  binomial("logit"), data = final_df)%>%  
  tbl_regression(exponentiate = TRUE,
                 show_single_row="treatment_cat1",
                 include = c("treatment_cat1"))
```


```{r, warning =F}
adjusted_hct_mediator_model <- glm(rc_transfusion_titans ~ treatment_cat1+ hct_week1_peak +GA_weeks_and_days_integer + multiple,  family =  binomial("logit"), data = final_df) %>%   
  tbl_regression(exponentiate = TRUE,
                 show_single_row="treatment_cat1",
                 include = c("treatment_cat1"))
```


```{r, warning =F}
adjusted_severity_of_illness_mediator_model <- glm(rc_transfusion_titans ~ treatment_cat1+ cumalitive_blood_vol_sampled_scaled +arterial_lines+mech_vent_combined +GA_weeks_and_days_integer + multiple,  family =  binomial("logit"), data = final_df) %>%   tbl_regression(exponentiate = TRUE,
                 show_single_row="treatment_cat1",
                 include = c("treatment_cat1"))
```


```{r, warning =F}
adjustedfullmodel <- glm(rc_transfusion_titans ~ treatment_cat1+ hct_week1_peak+cumalitive_blood_vol_sampled_scaled +arterial_lines+mech_vent_combined +GA_weeks_and_days_integer + multiple,  family =  binomial("logit"), data = final_df) %>%   tbl_regression(exponentiate = TRUE,
                 show_single_row="treatment_cat1",
                 include = c("treatment_cat1"))
```


```{r, warning =F}
models.a<-tbl_stack(list(unadjustedmodel, 
                         adjusted_basedlinedmodel, 
                         adjusted_hct_mediator_model,
                         adjusted_severity_of_illness_mediator_model,
                         adjustedfullmodel))

models.a$table_body %>%
  mutate(name = case_when(tbl_id1==1 ~ "Unadjusted", 
                          tbl_id1==2 ~ "Adjusted for baseline covariates" , 
                          tbl_id1==3 ~ "Hct Mediation adjusted",
                          tbl_id1==4 ~ "Severity of illness Mediation adjusted",
                          tbl_id1==5 ~ "Global model",) %>%
           as.factor()) %>%
  ggplot(aes(y=fct_reorder(name, -tbl_id1), x=estimate)) +
  geom_point(size=3)+
  geom_errorbar(aes(xmax=conf.high, xmin=conf.low), size=0.5, width=0.1) +
  labs(x="Odds ratio (95% CI)", y="")+
  geom_vline(aes(xintercept= 1), linetype="dotted")+
  coord_trans(x = scales:::log_trans(base = exp(1))) +
  theme_minimal() -> plot.a

plot.a
```

# Medflex (sequential approach)

https://academic.oup.com/ije/article/47/3/829/4829681?login=true 

kristy paper
https://academic.oup.com/ejendo/article/189/1/50/7219871?login=true
https://github.com/kristyrobledo/T4DM_mediation_paper

instruction paper
https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-022-01764-w

(sup materials)
https://static-content.springer.com/esm/art%3A10.1186%2Fs12874-022-01764-w/MediaObjects/12874_2022_1764_MOESM1_ESM.pdf

## Outcome: transufsion any (y/n)

### Mediation of Hct peak (M1 only) 


```{r, warning =F}
final_df1 <- final_df %>% filter(!is.na(rc_transfusion_titans) &
                                  !is.na(hct_week1_peak)&
                                   !is.na(GA_weeks_and_days_integer)&
                                  !is.na(multiple))


impData <- medflex::neImpute(rc_transfusion_titans ~ factor(treatment_cat1) +
                     hct_week1_peak +
                    GA_weeks_and_days_integer +
                    multiple,
                    family = binomial("logit"), nMed = 1, data = final_df1)


#head(impData)
neMod.extra.cont <- medflex::neModel(rc_transfusion_titans ~ treatment_cat10+
                              treatment_cat11 +
                    GA_weeks_and_days_integer +
                    multiple,
                  family = binomial("logit"),
                  expData = impData,
                  se = "robust")


#summary(neMod.extra.cont)

cont.extra<-data.frame(est = neMod.extra.cont$neModelFit$coefficients, confint(neMod.extra.cont))

lht <- medflex::neLht(neMod.extra.cont, linfct = c("treatment_cat101 = 0", 
                                          "treatment_cat111  = 0", 
                                          "treatment_cat101 + treatment_cat111  = 0"))

t<-summary(lht)

cat.extra<-data.frame(est = exp(t$coefficients[,1]), exp(confint(lht)))

#t


#pte.direct.cat<-(t$coefficients[1,1]/t$coefficients[3,1])*100
pte.indirect.cat<-(t$coefficients[2,1]/t$coefficients[3,1])*100
#pte.indirect.cat


cat.extra %>%
  mutate(name = c("Direct effect", "Indirect effect", "Total effect")) %>%
  ggplot(aes(y=fct_rev(name), x=est)) +
  geom_point(size=3)+
  geom_errorbar(aes(xmax=X95..UCL, xmin=X95..LCL), size=0.5, width=0.1) +
  labs(x="Odds ratio (95% CI)", y="")+
  geom_vline(aes(xintercept= 1), linetype="dotted")+
  coord_trans(x = scales:::log_trans(base = exp(1))) +
  theme_minimal() -> plot.a2

plot.a2
```


```{r}
#Formulat for table
cat.extra <- cat.extra %>%
  mutate(Effect = c("Direct effect", "Indirect effect", "Total effect")) %>% relocate(Effect) %>% dplyr::rename(
    "Estimate"= est,
    "95%_CI_L"= `X95..LCL`,
   "95%_CI_U"= `X95..UCL`,)
row.names(cat.extra) <- NULL





M1.model_transfusion_any <- cat.extra


M1.model_transfusion_any #%>% kable() %>% kable_styling()


prop_mediated_M1.model_trans_any <-  round((t$coef[3]/(t$coef[2]+t$coef[3])*100),1)
```


```{r}
print(paste0("Proportion mediated: ",prop_mediated_M1.model_trans_any , "%"))


```




## Joint model including both Hct (M1) and Severity of illness (M2)



```{r}
final_df1 <- final_df %>% filter(!is.na(hct_week1_peak)&
                                  !is.na(cumalitive_blood_vol_sampled)&
                                  !is.na(arterial_lines)&
                                  !is.na(mech_vent_combined)&
                                  !is.na(GA_weeks_and_days_integer)&
                                  !is.na(multiple))


#==========================================================================================================
impData <- medflex::neImpute(rc_transfusion_titans ~ factor(treatment_cat1) +
                     hct_week1_peak +cumalitive_blood_vol_sampled +  arterial_lines+ mech_vent_combined +
                    GA_weeks_and_days_integer +
                    multiple,
                    family = binomial("logit"), nMed = 4, data = final_df1)


#head(impData)
neMod.extra.cont <- medflex::neModel(rc_transfusion_titans ~ treatment_cat10+
                              treatment_cat11+
                    GA_weeks_and_days_integer +
                    multiple,
                  family = binomial("logit"),
                  expData = impData,
                  se = "robust")


#summary(neMod.extra.cont)

cont.extra<-data.frame(est = neMod.extra.cont$neModelFit$coefficients, confint(neMod.extra.cont))

lht <- medflex::neLht(neMod.extra.cont, linfct = c("treatment_cat101 = 0", 
                                          "treatment_cat111  = 0", 
                                          "treatment_cat101 + treatment_cat111  = 0"))

t<-summary(lht)

cat.extra<-data.frame(est = exp(t$coefficients[,1]), exp(confint(lht)))

#t


#pte.direct.cat<-(t$coefficients[1,1]/t$coefficients[3,1])*100
pte.indirect.cat<-(t$coefficients[2,1]/t$coefficients[3,1])*100
#pte.indirect.cat


cat.extra %>%
  mutate(name = c("Direct effect", "Indirect effect", "Total effect")) %>%
  ggplot(aes(y=fct_rev(name), x=est)) +
  geom_point(size=3)+
  geom_errorbar(aes(xmax=X95..UCL, xmin=X95..LCL), size=0.5, width=0.1) +
  labs(x="Odds ratio (95% CI)", y="")+
  geom_vline(aes(xintercept= 1), linetype="dotted")+
  coord_trans(x = scales:::log_trans(base = exp(1))) +
  theme_minimal() -> plot.a2

plot.a2
```


```{r}
#code to table results
cat.extra <- cat.extra %>%
  mutate(Effect = c("Direct effect", "Indirect effect", "Total effect")) %>% relocate(Effect) %>% dplyr::rename(
    "Estimate"= est,
    "95%_CI_L"= `X95..LCL`,
   "95%_CI_U"= `X95..UCL`)
row.names(cat.extra) <- NULL

#cat.extra


joint.M1M2.model_transfusion_any <- cat.extra
joint.M1M2.model_transfusion_any# %>% kable() %>% kable_styling()

prop_mediated_joint_model_trans_any <-  round((t$coef[3]/(t$coef[2]+t$coef[3])*100),1)
```


```{r}
paste0("Proportion mediated: ",prop_mediated_joint_model_trans_any, "%")


```



```{r, eval=F, include=F}
final_df1 <- final_df %>% filter(!is.na(hct_week1_peak)&
                                  !is.na(cumalitive_blood_vol_sampled)&
                                  !is.na(arterial_lines)&
                                  !is.na(mech_vent_combined)&
                                  !is.na(GA_weeks_and_days_integer)&
                                  !is.na(multiple))






impdata<-neImpute(rc_transfusion_titans ~ as.factor(treatment_cat1)+ hct_week1_peak*cumalitive_blood_vol_sampled*arterial_lines*mech_vent_combined +GA_weeks_and_days_integer + multiple,  family =  binomial("logit"), data = final_df1, nMed=4)
#head(impdata)
nemod=neModel(rc_transfusion_titans ~treatment_cat10+ treatment_cat11+ GA_weeks_and_days_integer +  multiple, family =  binomial("logit"), expData = impdata,se="bootstrap",nBoot = 1000)
summary(nemod)
lht<-neLht(nemod,linfct=c("treatment_cat101 = 0",
                                          "treatment_cat111  = 0",
                                          "treatment_cat101 + treatment_cat111  = 0"))
exp(cbind(coef(lht),confint(lht,type="perc"))) #confint with bootstrap
```





# Table of results

```{r}

M1_model_results <- bind_rows(list("M1 only: transfusion_any"=M1.model_transfusion_any), .id="id")
M1_model_results <- M1_model_results %>% mutate(across(c(Estimate, `95%_CI_L`, `95%_CI_U`), round, digits=2) )
M1_model_results$est_for_table <- paste0(M1_model_results$Estimate, " (", M1_model_results$`95%_CI_L`,"-", M1_model_results$`95%_CI_U`,")")

M1_model_results <- M1_model_results %>% select(id, Effect, est_for_table)


M1M2_joint_model_results <- bind_rows(list("M1M2 joint model: transfusion_any"=joint.M1M2.model_transfusion_any), .id="id")
M1M2_joint_model_results <- M1M2_joint_model_results %>% mutate(across(c(Estimate, `95%_CI_L`, `95%_CI_U`), round, digits=2) )
M1M2_joint_model_results$est_for_table <- paste0(M1M2_joint_model_results$Estimate, " (", M1M2_joint_model_results$`95%_CI_L`,"-", M1M2_joint_model_results$`95%_CI_U`,")")
M1M2_joint_model_results <- M1M2_joint_model_results %>% select(id, Effect, est_for_table)


combined_mediation_results <- cbind(M1_model_results, M1M2_joint_model_results)



combined_mediation_results %>% kable()



```

